Refinement of Thermostated Molecular Dynamics Using Backward Error Analysis

### 2019-02-26 rigid_bodies.tex
change tikz figure for eps

\usepackage[hyperindex,breaklinks,hidelinks,colorlinks,citecolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{natmove}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{tikz}
\usetikzlibrary{decorations.text}

@@

\begin{figure}
\begin{tikzpicture}
\draw[->, dashed]   (0,0.0) -- (7,0.0) node[right]{${\vt r}(t)$};
\node(0) at (0,0)[shape=circle,draw,double,scale=0.5]        {};
\node(1) at (2.2,0)[fill=black,shape=circle,draw,scale=0.5]        {};
\node(2) at (4.4,0)[fill=black,shape=circle,draw,scale=0.5]        {};
\node(3) at (6.6,0)[shape=circle,draw,double,scale=0.5] {};
\node(4) at (3.4,0.8)[] {$e^{\timestep \Liu{\tiny NVE}}$};
\node at (0.0,-0.2)[below]{${\vt r}_{t-2\timestep}$};
\node at (2.2,-0.2)[below]{${\vt r}_{t-\timestep}$};
\node at (4.4,-0.2)[below]{${\vt r}_{t}$};
\node at (6.6,-0.2)[below]{${\vt r}_{t+\timestep}$};
\path[->,thick] (1) edge [bend left=45] (2);
\path[dashed,->,thick,postaction={decorate,decoration={raise=2.5ex,text along path,text align=center,text={|\small|virtual step}}}] (0) to [bend left=45] (1);
\path[dashed,->,thick] (1) edge [bend right=45] (0);
\path[dashed,->,thick,postaction={decorate,decoration={raise=2.5ex,text along path,text align=center,text={|\small|virtual step}}}] (2) to [bend left=45] (3);
\path[dashed,->,thick] (2) edge [bend left=45] (3);
\end{tikzpicture}
\caption{Schematic representation of the executed steps in a single time-step of our refined NVT dynamics.
The solid arrow represents the NVE part of the integration, while the dotted arrows emphasize that forward and backward steps are carried out for positions and orientations with the only purpose of approximating their time-derivatives at point $t$.}
\label{fig:virtual steps}
\end{figure}

@@

\begin{figure}
	\includegraphics{NVE.eps}
	\caption{Specified Hamiltonian $\Ham{}$ (solid lines) and refined Hamiltonian $\refined{\Ham{}}$ (dotted lines) for different time-step sizes ($\timestep$) obtained from NVE MD simulations of 903 TIP3P \cite{Jorgensen_1983} water molecules using the numerical scheme of Eq.~\eqref{eq:splitting NVE} with the unsplit rotation method \cite{Silveira_2017}.}
	\label{fig:nve}
\end{figure}

@@

\begin{figure}
	\includegraphics{energy_drift.eps}
	\caption{Influence of the time-step size ($\timestep$) on the long-term drift rate ($R$) in: the extended energy ($\refined{H}$, black squares) and its refined Hamiltonian ($\refined{\Ham{}}$, green up triangles) part for our NVT method, as well as in the extended energy ($H$, red x) for the method due to Martyna \textit{et al}. \cite{Martyna_1996} and in the refined Hamiltonian ($\refined{\Ham{}}$, blue circles) for the numerical scheme of Eq.~\eqref{eq:splitting NVE} with unsplit solution of rotation.}
	\label{fig:energy_drift}
\end{figure}

@@

\begin{figure}
	\includegraphics{numerical_stability.eps}
	\caption{Time evolution of (a) the extended energy and (b) of each thermostat energy for the NVT method introduced by Kamberaj, Low, and Neal \cite{Kamberaj_2005}. In part (b), the negative values correspond to the thermostat coupled to the translational degrees of freedom.}
	\label{fig:num_stab}
\end{figure}

@@

\begin{figure*}
	\includegraphics{thermodynamic_properties.eps}
	\caption{Effect of the time-step size on (a) the estimated temperature, (b) the mean potential energy per molecule, (c) the mean internal virial per molecule, and (d) the constant-volume specific heat of water, as obtained from NVT-MD simulations of 903 TIP3P \cite{Jorgensen_1983} water molecules at $T=298~\mathrm{K}$ and $\rho=970~\mathrm{kg/m^3}$ and employing different numerical integration schemes.
	The NHC scheme ($\bullet$) corresponds to the Nos\'{e}-Hoover Chain integrator of Martyna \textit{et al}. \cite{Martyna_1996} extended to rigid bodies with analytically computed free rotations \cite{Silveira_2017}.
	The Refined NHC scheme ($\blacksquare$) is obtained by subjecting the preceding method to the refinement procedure described in Sec.~\ref{sec:refined_method}.
	The Refined NHC/NSQ scheme ($\blacktriangle$) differs from Refined NHC only in that the approximate NO-SQUISH \cite{Dullweber_1997, Miller_2002} solution is used for free rotations.
	The Refined CSVR scheme ($\times$) corresponds to the canonical stochastic velocity rescaling method \cite{Bussi_2007} subject to refinement as well.
	Finally, the KLN scheme ($\blacktriangledown$) is the double-thermostat method of Kamberaj, Low, and Neal \cite{Kamberaj_2005}.}
	\label{fig:properties}
\end{figure*}

@@

\begin{figure}
	\includegraphics{Free_Energy.eps}
	\caption{Difference in free energy per water molecule between the actually sampled and the target states as a function of the time-step size. The sampling was done by employing the Refined NHC scheme, while the target state is consistent with a distribution proportional to $e^{-\beta \Ham{}}$, where $\Ham{}$ corresponds to the specified Hamiltonian.}
	\label{fig:free_energy}
\end{figure}

@@

\begin{figure*}
	\includegraphics{energy_partition.eps}
		\caption{Influence of the time-step size on the kinetic energy partition in NVT-MD simulations of 903 TIP3P \cite{Jorgensen_1983} water molecules.
		Part (a) contains results obtained from the KLN method \cite{Kamberaj_2005}.
		The results in Part (b) were obtained by using the NHC method \cite{Martyna_1996} extended for rigid bodies and without any refinement.
		In Part (c), they  were obtained from our Refined NHC \cite{Martyna_1996} method, with the kinetic energies computed at the target state via reweighting.
		Finally, Part (d) depict the refined kinetic energies at every sampled state resulting from the same simulations of Part (c).
		In all parts, we present the simulated translational ($\blacktriangledown$) and rotational ($\blacktriangle$) kinetic energies, as well as the expected translational ($\blacksquare$) and rotational ($\times$) values assuming equipartition.}
	\label{fig:energy_partition}
\end{figure*}

>>>>

\usepackage[hyperindex,breaklinks,hidelinks,colorlinks,citecolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{natmove}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{epstopdf}

@@

\begin{figure}
	\includegraphics{scheme}
\caption{Schematic representation of the executed steps in a single time-step of our refined NVT dynamics. The solid arrow represents the NVE part of the integration, while the dotted arrows emphasize that forward and backward steps are carried out for positions and orientations with the only purpose of approximating their time-derivatives at point $t$.}
\label{fig:virtual steps}
\end{figure}

@@

\begin{figure}
	\includegraphics{NVE}
	\caption{Specified Hamiltonian $\Ham{}$ (solid lines) and refined Hamiltonian $\refined{\Ham{}}$ (dotted lines) for different time-step sizes ($\timestep$) obtained from NVE MD simulations of 903 TIP3P \cite{Jorgensen_1983} water molecules using the numerical scheme of Eq.~\eqref{eq:splitting NVE} with the unsplit rotation method \cite{Silveira_2017}.}
	\label{fig:nve}
\end{figure}

@@

\begin{figure}
	\includegraphics{energy_drift}
	\caption{Influence of the time-step size ($\timestep$) on the long-term drift rate ($R$) in: the extended energy ($\refined{H}$, black squares) and its refined Hamiltonian ($\refined{\Ham{}}$, green up triangles) part for our NVT method, as well as in the extended energy ($H$, red x) for the method due to Martyna \textit{et al}. \cite{Martyna_1996} and in the refined Hamiltonian ($\refined{\Ham{}}$, blue circles) for the numerical scheme of Eq.~\eqref{eq:splitting NVE} with unsplit solution of rotation.}
	\label{fig:energy_drift}
\end{figure}

@@

\begin{figure}
	\includegraphics{numerical_stability}
	\caption{Time evolution of (a) the extended energy and (b) of each thermostat energy for the NVT method introduced by Kamberaj, Low, and Neal \cite{Kamberaj_2005}. In part (b), the negative values correspond to the thermostat coupled to the translational degrees of freedom.}
	\label{fig:num_stab}
\end{figure}

@@

\begin{figure*}
	\includegraphics{thermodynamic_properties}
	\caption{Effect of the time-step size on (a) the estimated temperature, (b) the mean potential energy per molecule, (c) the mean internal virial per molecule, and (d) the constant-volume specific heat of water, as obtained from NVT-MD simulations of 903 TIP3P \cite{Jorgensen_1983} water molecules at $T=298~\mathrm{K}$ and $\rho=970~\mathrm{kg/m^3}$ and employing different numerical integration schemes.
	The NHC scheme ($\bullet$) corresponds to the Nos\'{e}-Hoover Chain integrator of Martyna \textit{et al}. \cite{Martyna_1996} extended to rigid bodies with analytically computed free rotations \cite{Silveira_2017}.
	The Refined NHC scheme ($\blacksquare$) is obtained by subjecting the preceding method to the refinement procedure described in Sec.~\ref{sec:refined_method}.
	The Refined NHC/NSQ scheme ($\blacktriangle$) differs from Refined NHC only in that the approximate NO-SQUISH \cite{Dullweber_1997, Miller_2002} solution is used for free rotations.
	The Refined CSVR scheme ($\times$) corresponds to the canonical stochastic velocity rescaling method \cite{Bussi_2007} subject to refinement as well.
	Finally, the KLN scheme ($\blacktriangledown$) is the double-thermostat method of Kamberaj, Low, and Neal \cite{Kamberaj_2005}.}
	\label{fig:properties}
\end{figure*}

@@

\begin{figure}
	\includegraphics{Free_Energy}
	\caption{Difference in free energy per water molecule between the actually sampled and the target states as a function of the time-step size. The sampling was done by employing the Refined NHC scheme, while the target state is consistent with a distribution proportional to $e^{-\beta \Ham{}}$, where $\Ham{}$ corresponds to the specified Hamiltonian.}
	\label{fig:free_energy}
\end{figure}

@@

\begin{figure*}
	\includegraphics{energy_partition}
		\caption{Influence of the time-step size on the kinetic energy partition in NVT-MD simulations of 903 TIP3P \cite{Jorgensen_1983} water molecules.
		Part (a) contains results obtained from the KLN method \cite{Kamberaj_2005}.
		The results in Part (b) were obtained by using the NHC method \cite{Martyna_1996} extended for rigid bodies and without any refinement.
		In Part (c), they  were obtained from our Refined NHC \cite{Martyna_1996} method, with the kinetic energies computed at the target state via reweighting.
		Finally, Part (d) depict the refined kinetic energies at every sampled state resulting from the same simulations of Part (c).
		In all parts, we present the simulated translational ($\blacktriangledown$) and rotational ($\blacktriangle$) kinetic energies, as well as the expected translational ($\blacksquare$) and rotational ($\times$) values assuming equipartition.}
	\label{fig:energy_partition}
\end{figure*}
